---
title: "Colour interpolation"
author: "Dr Annie Quinn (annievisserquinn@gmail.com)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 3
    theme: flatly
    highlight: breezedark
  pdf_document:
    toc: yes
    toc_depth: '3'
vignette: |
  %\VignetteIndexEntry{Intro to colR} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(colR)
library(viridis)
library(khroma)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, out.height = "75%", out.width = "75%", cache = TRUE)

myCols <- c("#00429d", "#ffffe0")
l1_lab <- colRpalLinear(myCols, nSteps = 5)

theme_lab <- function() { 
  theme_minimal() +
  theme(strip.placement = "outside",
        axis.title.x = element_blank(),
        panel.spacing = unit(0.5, "cm"),
        legend.position = "none", 
        panel.background = element_blank(), 
        panel.border = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0.5)) 
}
```

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri(file.path("../colR.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; max-width:20%')
```

Two or more colours can be interpolated to create colour palettes in colR:

- `colRpalLinear(hex, nSteps, colourSpace)` applies linear interpolation 

- `colRpalBezier(hex, nSteps, colourSpace)` applies Bezier interpolation, featuring smoother colour transitions

- `colRpalL(hex, colourSpace)` is a lightness smoother for Bezier interpolated colour palettes

<br>

The input arguments are:

- `hex`, a vector of at least 2 hexidecimal colours.
- `nSteps`, the number of colours to output.
- `colourSpace`, the name of the colour space for interpolation. Via the `farver` package, **colR supports twelve 3-dimensional colour spaces**; check out `?farver::decode_colour()`. The colR default is "lab" (CIElab).
- `col`, the position of lightness, from 1 to 3. The default is 1.

<br>

## 1. Examples in the (default) CIELAB colour space

Let's create colour palettes from the 3 colours below. **A vignette with technical details is under development.**

```{r, echo = FALSE, out.height = "75%", out.width = "75%"}
myCols <- c("#F3F3F6", "#DC426A", "#2D42A6")
myCols %>% colRplot
```

<br>

### 1.1 Creating the colour palettes

```{r}
palLinear <- colRpalLinear(myCols, nSteps = 9)
palBezier <- colRpalBezier(myCols, nSteps = 9)
palBezierLight <- palBezier %>% colRpalL
```

```{r}
tibble(palLinear = palLinear, 
       palBezier = palBezier,
       palBezierLight = palBezierLight) %>%
  colRtidy %>%
  colRplot
```

<br>

### 1.2 Colour palettes with cvd

```{r}
(palLinear %>% colRcvd %>% colRtidy %>% colRplot) + labs(title = "palLinear")
(palBezier %>% colRcvd %>% colRtidy %>% colRplot) + labs(title = "palBezier")
(palBezierLight %>% colRcvd %>% colRtidy %>% colRplot) + labs(title = "palBezierLight")
```

<br>

### 1.3 Applying the colour palettes to the diamonds dataset

```{r}
dt <- lapply(c("palLinear", "palBezier", "palBezierLight"), function(x) { 
  diamonds %>% 
    select(carat, price, clarity) %>% 
    mutate(pal = x, order = 1:n()) 
  }) %>% 
  bind_rows() %>%
  mutate(pal = factor(pal, levels = c("palLinear", "palBezier", "palBezierLight"))) %>%
  arrange(pal, clarity) %>%
  mutate(colour = paste0(pal, clarity) %>% fct_inorder) %>%
  arrange(order)
hex <- c(rev(palLinear[-1]), rev(palBezier[-1]), rev(palBezierLight[-1]))
```
```{r}
p1 <- 
  dt %>%
  ggplot(aes(x = carat, y = price, colour = colour)) + 
  geom_point(alpha = 0.75, size = 2.5) + 
  theme_minimal() +
  theme(legend.position = "none") +
    facet_wrap(~pal) +
    scale_colour_manual(values = hex)
p1
```
```{r}
p2 <- 
  dt %>%
  ggplot(aes(price, fill = colour)) +
  geom_histogram(position = "fill", binwidth = 2500) +
  facet_wrap(~pal) +
  scale_fill_manual(values = hex) +
  theme_minimal() +
  theme(legend.position = "none")
p2
```

<br>

## 2. Examples in other colour spaces

```{r}
palLinear <- 
palBezier <- 
palBezierLight <- palBezier %>% colRpalL
```

```{r}
tibble(palLinear = colRpalLinear(myCols, nSteps = 9, colourSpace = "rgb"), 
       palBezier = colRpalBezier(myCols, nSteps = 9, colourSpace = "rgb")) %>%
  colRtidy %>%
  colRplot
tibble(palLinear = colRpalLinear(myCols, nSteps = 9, colourSpace = "hsl"), 
       palBezier = colRpalBezier(myCols, nSteps = 9, colourSpace = "hsl")) %>%
  mutate(palBezierLight = palBezier %>% colRpalL(colourSpace = "hsl", col = 1)) %>%
  colRtidy %>%
  colRplot
tibble(palLinear = colRpalLinear(myCols, nSteps = 9, colourSpace = "xyz"), 
       palBezier = colRpalBezier(myCols, nSteps = 9, colourSpace = "xyz")) %>%
  colRtidy %>%
  colRplot
tibble(palLinear = colRpalLinear(myCols, nSteps = 9, colourSpace = "cmy"), 
       palBezier = colRpalBezier(myCols, nSteps = 9, colourSpace = "cmy")) %>%
  colRtidy %>%
  colRplot
```
