---
title: "Colour interpolation"
author: "Dr Annie Quinn (annievisserquinn@gmail.com)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 3
    theme: flatly
    highlight: breezedark
  pdf_document:
    toc: yes
    toc_depth: '3'
vignette: |
  %\VignetteIndexEntry{Intro to colR} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(colR)
library(viridis)
library(khroma)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, out.height = "75%", out.width = "75%", cache = TRUE)

myCols <- c("#00429d", "#ffffe0")
l1_lab <- colRpalLinear(myCols, nSteps = 5)

theme_lab <- function() { 
  theme_minimal() +
  theme(strip.placement = "outside",
        axis.title.x = element_blank(),
        panel.spacing = unit(0.5, "cm"),
        legend.position = "none", 
        panel.background = element_blank(), 
        panel.border = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0.5)) 
}
```

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri(file.path("../colR.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; max-width:20%')
```

Two or more colours can be interpolated to create colour palettes in colR:

- `colRpalLinear(hex, nSteps, colourSpace)` applies linear interpolation 

- `colRpalBezier(hex, nSteps, colourSpace)` applies Bezier interpolation, featuring smoother colour transitions

- `colRpalL(hex, colourSpace)` is a lightness smoother for Bezier interpolated colour palettes

The input arguments are:

- `hex`, a vector of at least 2 hexidecimal colours
- `nSteps`, the number of colours to output
- `colourSpace`, the name of the colour space for interpolation. Via the `farver` package, **colR supports twelve 3-dimensional colour spaces**; check out `?farver::decode_colour()`. The colR default is "lab" (CIElab)

Let's create colour palettes from the 3 colours below. For technical details check out **XYZ SOMETHING**.

```{r, echo = FALSE, out.height = "50%", out.width = "50%"}
myCols <- c("#F3F3F6", "#DC426A", "#2D42A6")
myCols %>% colRplot
```

## Examples in the (default) CIELAB colour space

```{r}
palLinear <- colRpalLinear(myCols, nSteps = 9)
palBezier <- colRpalBezier(myCols, nSteps = 9)
palBezierLight <- palBezier %>% colRpalL
```
```{r}
palLinear
palBezier
palBezierLight
```
What does the palette look like in the CIElab colour space?

```{r, echo = FALSE, out.height = "150%", out.width = "100%"}
l1_lab %>% 
  farver::decode_colour(to = "lab") %>% 
  as_tibble %>% 
  setNames(paste0(names(.), "*")) %>%
  mutate(hex = l1_lab, hex = hex %>% fct_inorder) %>%
  gather("facet", "value", -c(1,4)) %>%
  ggplot(aes(x = value, y = `l*`)) +
  geom_path(colour = "lightgrey", size = 1, linetype = 2) +
  geom_point(aes(fill = hex), size = 5, shape = 21, colour = "slategrey") +
  geom_text(aes(label = hex), size = 3, colour = "slategrey", hjust = 0, nudge_x = 15, nudge_y = -1) +
  facet_grid(~facet, switch = "x") +
  theme_lab() +
  scale_fill_manual(values = l1_lab) +
  scale_x_continuous(limits = c(-180,180), breaks = c(-180,-90,0,90,180), minor_breaks = NULL, expand = c(0,0)) + 
  scale_y_continuous(limits = c(0,100), breaks = c(0,50,100), minor_breaks = NULL, expand = c(0,0))
```


### Bezier interpolation
Splines create smooth curves from irregular data points, like a colours. The function `colRpalBezier()` uses Bezier interpolation to create colour palettes with smoother transitions. The inputs are the same as `colRpalLinear()`, hex, nSteps and colourSpace.

```{r}
l2_lab <- myCols %>% colRpalBezier(nSteps = 9)
l2_lab %>% colRplot
```

What does the palette look like in the CIElab colour space? The palette looks much smoother and colours are almost equidistant

```{r, echo = FALSE, out.height = "150%", out.width = "100%"}
l2_lab %>% 
  farver::decode_colour(to = "lab") %>% 
  as_tibble %>% 
  setNames(paste0(names(.), "*")) %>%
  mutate(hex = l2_lab, hex = hex %>% fct_inorder) %>%
  gather("facet", "value", -c(1,4)) %>%
  ggplot(aes(x = value, y = `l*`)) +
  geom_path(colour = "lightgrey", size = 1, linetype = 2) +
  geom_point(aes(fill = hex), size = 5, shape = 21, colour = "slategrey") +
  geom_text(aes(label = hex), size = 3, colour = "slategrey", hjust = 0, nudge_x = 15, nudge_y = -1) +
  facet_grid(~facet, switch = "x") +
  theme_lab() +
  scale_fill_manual(values = l2_lab) +
  scale_x_continuous(limits = c(-128,127), breaks = c(-128,0,127), minor_breaks = NULL, expand = c(0,0)) + 
  scale_y_continuous(limits = c(0,100), breaks = c(0,50,100), minor_breaks = NULL, expand = c(0,0))
```

### Bezier interpolation wiht lightness adjustment

Bezier interpolation can lead to an unequal transition in lightness between colours. This can be corrected using the function `colRpalL()` with two input arguments:

- hex, a vector of Bezier interpolated hexidecimal colours
- colourSpace, the name of the colour space for interpolation (see above); the default is "lab" (CIElab)

```{r}
l3_lab <- l2_lab %>% colRpalL
l3_lab %>% colRplot
```
What does the palette look like in the CIElab colour space? Notice that the colours spacing is now equidistant along the curve

```{r, echo = FALSE, out.height = "150%", out.width = "100%"}
l3_lab %>% 
  farver::decode_colour(to = "lab") %>% 
  as_tibble %>% 
  setNames(paste0(names(.), "*")) %>%
  mutate(hex = l2_lab, hex = hex %>% fct_inorder) %>%
  gather("facet", "value", -c(1,4)) %>%
  ggplot(aes(x = value, y = `l*`)) +
  geom_path(colour = "lightgrey", size = 1, linetype = 2) +
  geom_point(aes(fill = hex), size = 5, shape = 21, colour = "slategrey") +
  geom_text(aes(label = hex), size = 3, colour = "slategrey", hjust = 0, nudge_x = 15, nudge_y = -1) +
  facet_grid(~facet, switch = "x") +
  theme_lab() +
  scale_fill_manual(values = l2_lab) +
  scale_x_continuous(limits = c(-180,180), breaks = c(-180,-90,0,90,180), minor_breaks = NULL, expand = c(0,0)) + 
  scale_y_continuous(limits = c(0,100), breaks = c(0,50,100), minor_breaks = NULL, expand = c(0,0))
```
```{r}
pacman::p_load(patchwork)
p <- 
  diamonds %>%
  ggplot(aes(x = carat, y = price, colour = clarity)) + 
  geom_point(alpha = 0.75, size = 2.5) + 
  theme_minimal() +
  theme(legend.position = "top") +
  guides(colour = guide_legend(title.position = "top", nrow = 3))
p

(p + scale_colour_manual(values = rev(l1_lab)) + labs(colour = "l1")) + 
  (p + scale_colour_manual(values = rev(l2_lab)) + labs(colour = "l2")) + 
  (p + scale_colour_manual(values = rev(l3_lab)) + labs(colour = "l3"))
```

```{r}
dt <- lapply(c("l1_lab", "l2_lab", "l3_lab"), function(x) { 
  diamonds %>% 
    select(carat, price, clarity) %>% 
    mutate(pal = x, order = 1:n()) 
  }) %>% 
  bind_rows() %>%
  arrange(pal, clarity) %>%
  mutate(colour = paste0(pal, clarity) %>% fct_inorder) %>%
  arrange(order)
hex <- c(rev(l1_lab[-1]), rev(l2_lab[-1]), rev(l3_lab[-1])); hex

#p <- 
  dt %>%
  ggplot(aes(x = carat, y = price, colour = colour)) + 
  geom_point(alpha = 0.75, size = 2.5) + 
  theme_minimal() +
  theme(legend.position = "none") +
    facet_wrap(~pal) +
    scale_colour_manual(values = hex)
#p
```

```{r}
dt %>%
  ggplot(aes(carat, fill = colour)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(~pal) +
  scale_fill_manual(values = hex) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = seq(0,6,1), minor_breaks = NULL)
```

```{r}
dt %>%
  ggplot(aes(price, fill = colour)) +
  geom_histogram(position = "fill") +
  facet_wrap(~pal) +
  scale_fill_manual(values = hex) +
  theme_minimal() +
  theme(legend.position = "none")
```
