---
title: "Intro to colR"
author: "Dr Annie Quinn (annievisserquinn@gmail.com)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 3
    theme: flatly
    highlight: breezedark
  pdf_document:
    toc: yes
    toc_depth: '3'
vignette: |
  %\VignetteIndexEntry{Intro to colR} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(colR)
library(viridis)
library(khroma)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, out.height = "75%", out.width = "75%")
```

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri(file.path("../colR.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; max-width:20%')
```

## 1. Getting started

- Use the `devtools` package to install `colR` from GitHub: `devtools::install_github("avisserquinn/colR")`
- Load `colR` in R using `library(colR)`

If using the package manager `pacman`:

- Install and update `colR` using `pacman::p_install_gh("avisserquinn/colR")`
- Load `colR` in R using `pacman::p_load(colR)`

*Note: colR is under development. Check for updates on each use.*

<br>

## 2. Functions

```{r eval = FALSE, include = TRUE}
library(colR)
```

For ease of use, each `colR` function has the prefix *colR*. Print all functions using `ls("package:colR")`:

```{r, echo = FALSE}
ls("package:colR")
```

<br>

### 2.1 Example 1 - One colour

Let's explore the functions `colRplot()` and `colRcvd()` using a single colour, `#38A9AC` (the main colour in the colR hex sticker).

```{r}
myCol <- "#38A9AC"
```

The function `colRplot()` uses ggplot to visualise the colour. The colour is overlaid with the corresponding hex code.

```{r}
myCol %>% colRplot
```

The function `colRcvd()` transforms colour(s) to simulate three types of colour vision deficiency (cvd):

- Deutanomaly
- Protanomaly
- Tritanomaly

For ease of use, only the first three characters are required. It is *not* case sensitive.

The following examples all have the same output, `r myCol %>% colRcvd(cvd = "deu")`, a hex colour in vector form.

```{r, eval = FALSE}
myCol %>% colRcvd(cvd = "deu")
myCol %>% colRcvd(cvd = "deutanomaly")
myCol %>% colRcvd(cvd = "Deutanomaly")
myCol %>% colRcvd(cvd = "deuteranopia")
```

Omit the cvd argument to simulate all three types of colour vision deficiency.

```{r}
myCol %>% colRcvd
```

Combine `colRcvd()` with `colRplot()` to visualise.

```{r, out.height = "75%", out.width = "75%"}
myCol %>% colRcvd %>% colRplot
```

<br> 

### 2.3 Example 2 - Two colours
Now let's look at a vector of multiple colours.

```{r}
myCol <- c("#2ECC71", "#F1C40F", "#E74C3C")
```

But what colours are they? We can easily visualise them using `colRplot()`.

```{r}
myCol %>% colRplot
```

As before, the colours are labelled with their hex code. On the left-hand side you can see the colour number, this is the colour index, or position, in the vector.

How does cvd change these colours?

```{r}
myCol %>% colRcvd
```

This time the output is a matrix, not a vector. Use `colRtidy()` to tidy the matrix for input to `colRplot()`.

```{r}
myCol%>% colRcvd %>% colRtidy 
myCol %>% colRcvd %>% colRtidy %>% colRplot
```

## 3. Example applications

### 3.1 Visualising colour palettes
```{r}
viridis::viridis(option = "A", n = 10) %>% colRplot
viridis::viridis(option = "B", n = 10) %>% colRplot
viridis::viridis(option = "C", n = 10) %>% colRplot
viridis::viridis(option = "D", n = 10) %>% colRplot
```

<br>

### 3.2 Finding a distinct colour colour palette

Check out the first two colour palettes from viridis. 

```{r}
viridis::viridis(option = "A", n = 10) %>%
  colRcvd %>%
  colRtidy %>%
  colRplot

viridis::viridis(option = "B", n = 10) %>%
  colRcvd %>%
  colRtidy %>%
  colRplot
```

The first colours are not very distinct! They're not well-suited for a qualitative palette.

```{r}
viridis::viridis(option = "D", n = 10) %>%
  colRcvd %>%
  colRtidy %>%
  colRplot
```

Option D is a much more distinct option.

<br>

### 3.3 Finding a colour blind friendly distinct colour palette

Let's check out how accessible Paul Tol's colour palettes are; we can do this via the khroma package, https://cran.r-project.org/web/packages/khroma/vignettes/tol.html. The bright, contrast, vibrant and muted colour schemes should be colour blind friendly.

```{r}
khroma::colour("bright")(7) %>% as.vector %>% colRcvd %>% colRtidy %>% colRplot
khroma::colour("contrast")(3) %>% as.vector %>% colRcvd %>% colRtidy %>% colRplot
khroma::colour("vibrant")(7) %>% as.vector %>% colRcvd %>% colRtidy %>% colRplot
khroma::colour("muted")(9) %>% as.vector %>% colRcvd %>% colRtidy %>% colRplot
```

Bright, vibrant and muted have quite a few grey tones that could be difficult to discern. Contrast looks like it would work well for a qualitative palette.

<br>

### 3.4 Finding a colour blind friendly diverging colour palette

Let's look at Paul Tol's colour palettes again to see which are distinct at both ends of the spectrum. 

```{r}
khroma::colour("sunset")(9) %>% as.vector %>% colRcvd %>% colRtidy %>% colRplot
khroma::colour("BuRd")(9) %>% as.vector %>% colRcvd %>% colRtidy %>% colRplot
khroma::colour("PRGn")(9) %>% as.vector %>% colRcvd %>% colRtidy %>% colRplot
```

Only Sunset is distinct at both ends - the other two palettes 

<br>

### 3.5 Finding a colour blind friendly continuous colour palette

Let's check out the rainbow colour palette. Are the colour changes uniform?

```{r}
khroma::colour("smooth rainbow")(25) %>% as.vector %>% colRcvd %>% colRtidy %>% colRplot
```
We can see that the colours vary from dark to light across the palette. Let's check out an alternative, YlOrBr. 

```{r}
khroma::colour("YlOrBr")(9) %>% as.vector %>% colRcvd %>% colRtidy %>% colRplot
```
